---
title: "Evaluating approaches to backcalculating cases counts by date of infection from cases counts by date of report"
output: github_document
---

**Authors:** EpiForecasts, CMMID Covid working group, Sebastian Funk


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 8, 
  fig.align = "center",
  cache = FALSE,
  fig.path = "figures/"
)
```

## Summary

* The current methods being used to backcalculate case counts by date of infection assume either a constant shift (typically by the mean/median of the delay from onset to report and the mean/median of the incubation period) or assume independence between cases and sample both the delay from onset to report and the incubation period. 

* Parametric approaches are problematic in that they make assumptions about the data generating process that may not be correct.

* Here we evaluate a non-parametric approach that is based on convoluting the known delays in comparison to our currently deployed independent sampling method.

* We find that the non-parametric approach is better able to reproduce simulated data and returns plausible results when used on reported Covid-19 cases from 4 countries.

* We also implement an in model approach to estimating the time-varying reproduction number but this work is at an early stage.

* These findings are preliminary and more methodological work is required along with comparisons to other approaches.

## Introduction


## Dependencies

```{r, include = FALSE}
# To install dependencies
# library(drat); drat:::add("epiforecasts")
# and install all packages using install.packages as normal
library(EpiNow)
library(NCoVUtils)
library(data.table)
library(ggplot2)
library(cowplot)
library(lubridate)
library(purrr)
library(rstan)
library(patchwork)
library(truncnorm)

## non-parameteric nowcasting function
source("nowcast.R")

## Stan model
model <- rstan::stan_model("nowcast.stan")

## Run in parallel when possible
future::plan("multisession", workers = round(future::availableCores() / 2))

## Logical to control testing on real data
run_countries <- TRUE
```

## Methods

### Model

* Overview: non-parameteric backcalculation assuming a median shift with independent gaussian noise.

* Orginal implementation here: https://github.com/jhellewell14/rtconfirm/blob/master/R/backcalc.R

* This implementation uses median shifted reported cases (smoothed using a rolling average over the width of the median generation interval) as a prior and then fits independent gaussian noise on top of this. For future cases (i.e with no data to shift to into the last reported case count is used).

* Daily reporting effects are included and constrained by weekly cases.

* Uncertainty in reporting delays and incubation periods is enforced by upweighting the prior for these parameters.

* We implement an in model time-varying reproduction number estimation step based on the renewal equation and the imputed infections. 

```{r}
model
```

## Analysis

### Simulate data

* Define a realistic basic reproduction number estimate that starts at 2, decreases linearly to 0.5, remains constant, increases linearly to 1.2, decreases sharply to 1,  again remains constant and finally increases linearly. Noise is added (2% standard deviation) for all values.

```{r}
## Define an initial rt vector 
rts <- c(rep(2, 20), (2 - 1:15 * 0.1), rep(0.5, 10), (0.5 + 1:7 * 0.1), rep(1.2, 20),
         rep(1, 10), (1 + 1:10 * 0.05))
## Add noise
##rts <- rts * rnorm(length(rts), 1, 0.02)

rts
```

* In order to simulate cases by date of infection from a reproduction number trace an estimate of the generation interval is required. We use the current default estimate for Covid-19 from `EpiNow`.

```{r}
## Use the mean default generation interval for covid
generation_time_pdf <- rowMeans(EpiNow::covid_generation_times)

generation_time_pdf
```

```{r}
incubation_period <- list(mean = EpiNow::covid_incubation_period[1, ]$mean,
                           mean_sd = EpiNow::covid_incubation_period[1, ]$mean_sd,
                           sd = EpiNow::covid_incubation_period[1, ]$sd,
                           sd_sd = EpiNow::covid_incubation_period[1, ]$sd_sd,
                           max = 30)
                    
reporting_delay <- list(mean = log(5),
                         mean_sd = log(1.1),
                         sd = log(2),
                         sd_sd = log(1.2),
                         max = 30)

generation_time <- list(mean = EpiNow::covid_generation_times_summary[1, ]$mean,
                        mean_sd = EpiNow::covid_generation_times_summary[1, ]$mean_sd,
                        sd = EpiNow::covid_generation_times_summary[1, ]$sd,
                        sd_sd = EpiNow::covid_generation_times_summary[1, ]$sd_sd,
                        max = 30)

## Sample a report delay as a lognormal - take 10 samples
delay_defs <- EpiNow::lognorm_dist_def(mean = reporting_delay$mean, mean_sd = reporting_delay$mean_sd,
                                       sd = reporting_delay$sd, sd_sd = reporting_delay$sd_sd,
                                       max_value = reporting_delay$max, samples = 1000)


## Sample a incubation period (again using the default for covid) - take 10 samples
incubation_defs <- EpiNow::lognorm_dist_def(mean = incubation_period$mean,
                                           mean_sd = incubation_period$mean_sd,
                                           sd = incubation_period$sd,
                                           sd_sd = incubation_period$sd_sd,
                                           max_value = 30, samples = 1000)

## Simulate cases with a decrease in reporting at weekends and an incease on Monday
## using a single sample of both distributions                                    
simulated_cases <- EpiNow::simulate_cases(rts, initial_cases = 10 , initial_date = as.Date("2020-03-01"),
                                          generation_interval = generation_time_pdf,
                                          delay_def = delay_defs[10, ],
                                          incubation_def = incubation_defs[10, ],
                                          reporting_effect = c(1.6, 1, 1, 1, 0.8, 0.4, 1))
simulated_cases
```
    
  
###  Compare approaches on simulated data

```{r}
## Extract simulated infections
simulated_reports <- simulated_cases[reference == "report"][, confirm := cases][, cases := NULL]

## Rt prior
rt_prior <- list(mean = 2.6, sd = 2)

## Reconstruction via backwards sampling
sampling_cases <- nowcast_pipeline(reported_cases = simulated_reports[, import_status := "local"], 
                                   target_date = max(simulated_reports$date),
                                   delay_defs = delay_defs, 
                                   incubation_defs = incubation_defs,
                                   nowcast_lag = 0, approx_delay = TRUE)

## Non-parameteric reconstruction
non_parametric_cases <- nowcast(simulated_reports,
                                family = "negbin",
                                incubation_period = incubation_period,
                                reporting_delay = reporting_delay,
                                generation_time = generation_time, 
                                estimate_rt = FALSE,
                                rt_prior = rt_prior,
                                cores = 4, chains = 4,
                                samples = 1000, warmup = 1000,
                                return_all = TRUE, model = model,
                                verbose = TRUE)
```

### Compare approaches on reported Covid-19 cases in Austria, the United Kingdom, United States of America and Russia

* Get data

```{r}
reported_cases <- NCoVUtils::get_ecdc_cases(countries = c("Austria", "United_Kingdom",
                                                          "United_States_of_America", "Russia"))
reported_cases <- NCoVUtils::format_ecdc_data(reported_cases)
reported_cases <- data.table::as.data.table(reported_cases)[, confirm := cases][, cases := NULL]
reported_cases <- reported_cases[date >= "2020-02-01"]
```

* Run backcalculation on each country in turn

```{r, message = FALSE, eval = run_countries}
countries <- c("Austria", "United Kingdom", "United States of America", "Russia")

results <- lapply(countries,
                  function(country) {
        message("Nowcasting using sampling for: ", country)                
        cases <- data.table::copy(reported_cases)[region %in% country]  
                                        
        ## Reconstruction via backwards sampling
        sampling_cases <- nowcast_pipeline(reported_cases = cases[, import_status := "local"], 
                                           target_date = max(cases$date),
                                           delay_defs = delay_defs,
                                           incubation_defs = incubation_defs,
                                           nowcast_lag = 0, approx_delay = TRUE)
        
        message("Non-parametric nowcasting for: ", country)
        ## Non-parametric reconstruction
        non_parametric_cases <- nowcast(cases,
                                        family = "negbin",
                                        incubation_period = incubation_period,
                                        reporting_delay = reporting_delay,
                                        generation_time = generation_time, 
                                        estimate_rt = FALSE,
                                        samples = 2000, warmup = 1000,
                                        cores = 4, chains = 4,
                                        return_all = TRUE, model = model)
        
        return(list(sampling_cases, non_parametric_cases))
                                       })

names(results) <- countries
```

## Results


### Simulated data

* Reporting effects.

```{r}
non_parametric_cases$day_of_week[, as.list(summary(value)), by = "wday"]
```

* Recover reporting delays

```{r}
data.table::rbindlist(list(
  non_parametric_cases$rep_mean[, .(parameter = "mean", mean = mean(value), sd = sd(value))],
  non_parametric_cases$rep_sd[, .(parameter = "sd", mean = mean(value), sd = sd(value))]
))
```

* Recover incubation period

```{r}
data.table::rbindlist(list(
  non_parametric_cases$inc_mean[, .(parameter = "mean", mean = mean(value), sd = sd(value))],
  non_parametric_cases$inc_sd[, .(parameter = "sd", mean = mean(value), sd = sd(value))]
))
```

* Prepare data for plotting

```{r}
simulated_cases <- simulated_cases[reference %in% c("infection", "report")][, median := cases][,
                                   type := ifelse(reference == "infection", 
                                                  "Simulated infections", 
                                                  "Simulated reported cases")][,
                                   `:=`(cases = NULL, reference = NULL)]

summarise_nowcasting_approaches <- function(sampling_cases, non_parametric_cases) {
  sampling_cases <- sampling_cases[type %in% "infection_upscaled"][,
               .(median = median(cases), bottom = quantile(cases, 0.025),
                 lower = quantile(cases, 0.25), upper = quantile(cases, 0.75),
                 top = quantile(cases, 0.975)), by = c("date", "type")][,
                 type := "Sampled"]

non_parametric_infections <- non_parametric_cases$infections[,
               .(median = median(value), bottom = quantile(value, 0.025),
                 lower = quantile(value, 0.25), upper = quantile(value, 0.75),
                 top = quantile(value, 0.975)), by = c("date")][,
                 type := "Non-parametric"]

non_parametric_prior_infections <- non_parametric_cases$prior_infections[sample == 1,
               .(median = value), by = c("date")][,
                 type := "Infections prior"]
out <- data.table::rbindlist(list(sampling_cases, non_parametric_infections,
                                  non_parametric_prior_infections), fill = TRUE)
return(out)
}

summarised_nowcasting_approaches <- summarise_nowcasting_approaches(sampling_cases, non_parametric_cases)

simulated_cases <- data.table::rbindlist(list(simulated_cases,
                                              summarised_nowcasting_approaches),
                                         fill = TRUE)
```



* Plot date. *Note: Here we have cut-off cases prior to the start of March. This truncates the long tale observed in the sampling approach.*

```{r, fig.width = 9, fig.height = 9, dpi = 400, message = FALSE, warning = FALSE} 
plot_data <- simulated_cases[date >= as.Date("2020-03-01")]

plot_obs <- ggplot2::ggplot(plot_data[type %in% "Simulated reported cases"], 
                           ggplot2::aes(x = date, col = type, fill = type)) +
  ggplot2::geom_col(ggplot2::aes(y = median), fill = "grey", col = "white",
                    show.legend = FALSE) +
  cowplot::theme_cowplot() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::scale_color_brewer(palette = "Dark2") +
  ggplot2::labs(y = "Reported cases", x = "")


plot_sims_data <- plot_data[!type %in% "Simulated reported cases"][,
                             type := factor(type, levels = c("Simulated infections",
                                                             "Sampled",
                                                             "Infections prior",
                                                             "Non-parametric"))]
plot <- ggplot2::ggplot(plot_sims_data, ggplot2::aes(x = date, col = type, fill = type)) +
 ggplot2::geom_linerange(data = plot_sims_data[!type %in% "Simulated infections"],
                          ggplot2::aes(ymin = bottom, ymax = top), 
                         alpha = 0.2, size = 1.5) +
 ggplot2::geom_linerange(data = plot_sims_data[!type %in% "Simulated infections"],
                         ggplot2::aes(ymin = lower, ymax = upper), 
                         alpha = 0.4, size = 1.5) +
 ggplot2::geom_line(data = plot_sims_data[type %in% c("Simulated infections", "Infections prior")],
                    ggplot2::aes(y = median), size = 1.1, alpha = 0.6) +
 ggplot2::geom_point(data = plot_sims_data[type %in%  c("Simulated infections", "Infections prior")],
                     ggplot2::aes(y = median), size = 1.1, alpha = 0.8, show.legend = FALSE) +
  cowplot::theme_cowplot() +
  ggplot2::theme(legend.position = "bottom") +
  ggplot2::scale_color_brewer(palette = "Dark2") +
  ggplot2::labs(y = "Cases by date of infection", x = "Date", col = "Type")

plot_obs + plot + patchwork::plot_layout(ncol = 1)
```

### Reported Covid-19 cases in the United Kingdom, United States of America and South Korea

* Explore reporting effects by country

```{r, eval = run_countries}
purrr::map(results, ~ .[[2]]$day_of_week[, as.list(summary(value)), by = "wday"])
```

* Prepare data for plotting

```{r, eval = run_countries}
summarised_cases <- data.table::copy(reported_cases)[, median := confirm][, type := "Reported Cases"]


summarised_results <- purrr::map2(results, names(results),
                                  ~ summarise_nowcasting_approaches(.x[[1]], .x[[2]])[,region := .y])

summarised_results <- data.table::rbindlist(summarised_results)
all_country_data <- data.table::rbindlist(list(summarised_cases, summarised_results), fill = TRUE)
all_country_data <- all_country_data[date >= "2020-03-01"]
```

* Plot data

```{r, eval = run_countries, fig.width = 12, fig.height = 9, dpi = 400, message = FALSE, warning = FALSE} 
plot <- ggplot2::ggplot(all_country_data, ggplot2::aes(x = date, col = type, fill = type)) +
  ggplot2::geom_col(data = all_country_data[type %in% "Reported Cases"],
                    ggplot2::aes(y = median), fill = "grey", col = "white") +
  ggplot2::geom_linerange(ggplot2::aes(ymin = bottom, ymax = top), 
                         alpha = 0.2, size = 1) +
 ggplot2::geom_linerange(ggplot2::aes(ymin = lower, ymax = upper), 
                         alpha = 0.4, size = 1) +
  cowplot::theme_cowplot() +
  ggplot2::theme(legend.position = "bottom") +
  ggplot2::scale_color_brewer(palette = "Dark2") +
  ggplot2::labs(y = "Cases", x = "Date", col = "Type") + 
  ggplot2::facet_wrap(~region, scales = "free_y")

plot
```

## Discussion

## References
